Seismic parameter estimation and the Canadian crust

### 2013-06-28 paper/draftBP.tex
paper

\documentclass[draft, 12pt]{article}
\usepackage[english]{babel}
\usepackage[final]{graphicx}
\usepackage{color}
\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.png}

@@

  Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod} provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations.

As regional studies exist which utilize similar methods and have corresponding parameter estimates for some of the stations used in this study it is possible to directly compare values to those previously published. A comprehensive study in the Hudson Bay region of the Canadian Shield (Thomson et. al., 2010) uses an approach similar to the Zhu and Kanamori method published estimates for $H$ and $R=\frac{V_P}{V_S}$ for 35 stations. 5 stations with $R$ estimates above 1.8 are removed due to the uncertainty surrounding these high values. There is strong correlation of 0.95 between crustal thickness values for both datasets (Figure~\ref{fig:thompsonCompH}). The velocity ratio data has a lower correlation of 0.5 (Figure~\ref{fig:thompsonCompR}). [Compare this value to uncertainty / deviation in the data? ].

Directly comparing $V_P$ estimates processed with the MB algorithm is not possible as this method has not been employed in publication. However, comparisons can be made to active source records for experiments within close proximity to a given seismic station (Figure~\ref{fig:activeComp}. The correlation between these datasets is 0.235. [This is low, explain here?]. Another method of determining the reliability of estimates is to check the $\frac{V_P}{V_S}$ ratio taken from the Kanamori approach to the $\frac{V_P}{V_S}$ value computed from the MB approach. These values, working with the same data, preprocessed using the same methods and tools should be equal. Selecting only those stations with an error of less than +/-0.05 $R$, 132 stations, we get a correlation of 0.57. Reducing the error tolerance and selecting stations with less than +/-0.01 $R$, leaving 31 of the cleanest stations, we get a correlation coefficient of 0.95.

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{thompsonComparisonH}
  \caption{Comparison of crustal thickness $H$ data from this study with data from Thomspson et. al. (2010). Data shows a Pearson correlation of 0.95}
  \label{fig:thompsonCompH}
\end{figure}

@@

@@

\subsection{Crustal Thickness}
 (MAY NOT INCLUDE)
\begin{itemize}
  \item Jump in crustal thickness over the fault as you go from the Slave to the Rae or Churchill Province.
  \item Trend in Southern Ontario as we move north and up the St Lawrence Seaway.
  \item Appears to be secular variation in the North Eastern Churchill. As we move North and East we have thickening of continental crust.
\end{itemize}

\subsection{Vp/Vs}
\begin{itemize}
  \item reference Figure~\ref{fig:VpVsMap}
  \item For perspective, the value for Quartzite is 1.65 and Basalt is 1.85.
  \item North East of the Slave province, has cluster of low Vp/Vs values in a semi circle of high values.
  \item Appears to be an shift throughout the Superior province from high Vp/Vs in the South to low Vp/Vs in the North East.
  \item Compare to Crust 2.0 overlay (Figure~\ref{fig:VpVsMapCrust}).  Average value for Churchill province Vp/Vs is 1.73 while average value computed from Crust 2.0 is 1.77.
\end{itemize}

>>>>

\documentclass[draft, 12pt]{article}
\usepackage[english]{babel}
\usepackage[final]{graphicx}
\usepackage{multirow}
\usepackage{color}
\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.png}

@@

  Several published regional studies utilizing the methods outlined in section \ref{section:VpVsMethod} provide parameter estimates for $R$ and $H$ for regionally grouped subsets of Canadian seismic stations. Comparisons between estimates for particular seismic stations are made for those values which have a corresponding standard error in $R$ of less than 0.06. For each study under comparison, data is reprocessed to use the $V_P$ chosen by the study authors and both correlation and mean difference are calculated (Table \ref{table:comparison}).

  A comprehensive study in the Hudson Bay region of the Canadian Shield [Thomson et. al., 2010] employs data from 35 stations, of which, 27 stations share an error acceptable for comparison in both datasets. There is strong correlation, 0.97, between crustal thickness values (Fig \ref{fig:thompsonCompH}). The velocity ratio data has a lower correlation of 0.49 (Fig \ref{fig:thompsonCompR}) and a mean difference of 0.02. As indicated in section \ref{section:VpVsMethod} there is less resolution in $R$ so the lower correlation for this parameter is not unexpected.

 Another study conducted in the Grenville Orogen [Eaton et al., 2006] utilizes data from 29 stations, with two stations being excluded due to low quality and high error. Again, there is strong correlation in $H$, 0.89, and lower correlation in $R$, 0.56 (Figs \ref{fig:eatonCompH}, \ref{fig:eatonCompR}). The mean difference in $R$ between the two datasets is 0.04.

\begin{table}
  \begin{tabular}{ l l l l l l }
    \cline{3-6}
    & & \multicolumn{2}{ c }{Correlation} & \multicolumn{2}{ c }{Mean Difference} \\
    \hline
    Study Authors & Num. of Stns & $H$ & $R$ & $H$(km) & $R$ \\
    \hline
    Thompson et. al. (2010)   & 0.97 & 0.49 & 3.36 & 0.024 \\
    Eaton et. al. (2006)      & 0.90 & 0.61 & 2.65 & 0.04  \\
    Darbyshire et. al. (2007) & 0.95 & 0.43 & 4.16 & 0.39  \\
    \hline
  \end{tabular}
  \caption{Comparison of $R$ and $H$ estimates with three published studies}
\label{table:comparison}

\end{table}

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{thompsonComparisonH}
  \caption{Comparison of crustal thickness $H$ data from this study with data from Thompson et. al. (2010). Data shows a Pearson correlation of 0.95}
  \label{fig:thompsonCompH}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{eatonComparisonH}
  \caption{Comparison of crustal thickness $H$ data from this study with from Eaton et. al. (2010). Data shows a Pearson correlation of 0.95}
  \label{fig:eatonCompH}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{eatonComparisonR}
  \caption{Comparison of $V_P / V_R$ data from this study with data from Eaton et. al. (2010). Data shows a Pearson correlation of 0.5}
  \label{fig:eatonCompR}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{darbyshireComparisonH}
  \caption{Comparison of crustal thickness $H$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.95}
  \label{fig:darbyshireCompH}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{darbyshireComparisonR}
  \caption{Comparison of $V_P / V_R$ data from this study with data from Darbyshire et. al. (2010). Data shows a Pearson correlation of 0.5}
  \label{fig:darbyshireCompR}
\end{figure}

@@

\subsection{Regional Bulk Crustal Parameters}
(MAY NOT INCLUDE)
\begin{itemize}
  \item Jump in crustal thickness over the fault as you go from the Slave to the Rae or Churchill Province.
  \item Trend in Southern Ontario as we move north and up the St Lawrence Seaway.
  \item Appears to be secular variation in the North Eastern Churchill. As we move North and East we have thickening of continental crust.
\end{itemize}
